Development and Implementation of an ANN based flow law for Numerical Simulations of Thermo-mechanical Processes at High Temperatures in FEM Software

### 2022-12-15 Article.tex
Update of whole paper after reading

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.
\usepackage{gensymb}
\usepackage{amsmath}
\usepackage{accents}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{subcaption}
\DeclareRobustCommand{\w}{\mbox{\large\ensuremath{\mathsf{w}}}}
\DeclareRobustCommand{\A}{\mbox{\large\ensuremath{\mathsf{A}}}}
\DeclareRobustCommand{\B}{\mbox{\large\ensuremath{\mathsf{B}}}}
\DeclareRobustCommand{\C}{\mbox{\large\ensuremath{\mathsf{C}}}}
\DeclareRobustCommand{\Dev}{\mbox{\Large\ensuremath{\mathsf{s}}}}
\DeclareRobustCommand{\Sig}{\mbox{\Large\ensuremath{\sigma}}}
\DeclareRobustCommand{\dotp}{\boldsymbol{\cdot}}
\DeclareRobustCommand{\ccirc}{\kern0.5ex\vcenter{\hbox{$\scriptstyle\circ$}}\kern0.5ex}
\DeclareRobustCommand{\e}[1]{{\rm e}^{#1}}
\DeclareRobustCommand{\lay}[1]{^{(#1)}}
\DeclareRobustCommand{\mdot}[1]{\accentset{\mbox{\bfseries .}}{#1}}
\DeclareRobustCommand{\ie}{\emph{i.e.}\@\xspace}
\DeclareRobustCommand{\eal}{et \emph{al.}\@\xspace}
\DeclareRobustCommand{\eg}{e.g.,\@\xspace}
\DeclareRobustCommand{\RMSE}{\text{E}_\text{RMS}}
\DeclareRobustCommand{\AARE}{\text{E}_\text{AAR}}
\DeclareRobustCommand{\R}{\text{R}}
\DeclareRobustCommand{\ps}{\text{s}^{-1}}
\DeclareRobustCommand{\mr}[2]{\multirow{#1}{*}{#2}}
\DeclareRobustCommand{\MPa}{\text{MPa}}

@@

Numerical methods for simulating the behavior of structures subjected to high thermomechanical loads, as in the case of high-temperature forming of metallic materials, are generally based on the use of commercial Finite Element (FE) codes such as Abaqus or laboratory codes such as DynELA \cite{Pantale-2004}, which has been developed in our Laboratory for many years.
These simulation codes are based on two types of equations: conservation equations and constitutive equations.
If the first equations are generally well established on the basis of physics and mechanics, it is not the same for the second type of equations: the constitutive equations.
Thus, in a general way, the conservation equations concern the fundamental principles of physics such as the mass conservation law, the momentum law (fundamental equation) and the energy law (declined in the form of the first and second principles of thermodynamics).
By themselves, these laws are not sufficient to describe the behavior of a material or a structure subjected to thermomechanical stresses, because the nature of the material's behavior, generally translated by means of the behavior laws, is not included in the system previously proposed.
Therefore, for each type of material, it is necessary to define behavior laws whose formulation is based on observation, in order to describe the behavior of this material under external stresses.
The quality and the accuracy of the results of any numerical simulation depend on the choice of these behavior laws and on the ability of the user to identify the coefficients of these behavior laws for a given material by performing experiments under conditions close to those encountered during the real stress of the structure in service that one wishes to design.
Depending on the nature of the stresses, these tests are based on quasi-static or dynamic tensile or compression tests, tests on thermomechanical simulators such as Gleeble or impact tests using gas launchers or Hopkinson bars.

Generally speaking, in the context of thermomechanical simulation of forming processes, these behavior laws define the dependence of the flow stress of the material $\sigma^y$ as a function of the three input variables which are the plastic strain $\varepsilon^p$, the strain rate $\mdot\varepsilon$ and the temperature $T$ of the material.
These laws, due to the nature of materials and the phenomena involved (work hardening, movement of dislocations, structural hardening, phase transformations, ...) are highly non-linear and their validity is restricted to a certain range of strains, strain rates and temperatures.
From the observations made, we can define two main classes of behavior laws: the flow laws based on physics and the empirical flow laws.
From the mechanics of continuous media and experimental tests and depending on the nature of the materials used, a number of flow models have been developed in the past, including: the Johnson-Cook flow law \cite{Johnson-1983}, the Zerilli Armstrong flow law \cite{Zerilli-1987} and their respective derived forms \cite{Rule-1998, Lin-2010, Li-2013, Zhang-2015, Zhou-2019}, the Hansel-Spittle \cite{Hensel-1978} or the Arrhenius \cite{Jonas-1969} flow laws, to name only a few of the most widely used in the context of the shaping of metallic materials at high temperature.
Once the choice has been made concerning the type of flow law to be used for a given material, it is then necessary, from a set of experimental tests carried out in the laboratory under conditions close to those of the structure in service, to identify the parameters of these flow laws by machine learning methods based on approaches of minimization of the calculated experiment.

The main problem that the researchers are confronted with after the phase of realization of the experimental tests concerns the choice of the flow law to use according to the observations made on these test results.
This choice of flow law is also restricted by the nature of the FE code used.
Thus, a user of the Abaqus FE code, will turn more particularly to a flow law of the Johnson-Cook \cite{Johnson-1983} type insofar as it is natively implemented in this software.
The choice of another form of flow law (Zerilli-Armstrong, or Arrhenius for example) obliges the user to program himself the computation of the flow stress $\sigma^y$ of the material by means of a VUMAT routine in FORTRAN as proposed by Gao \eal \cite{Gao-2007-FRT}, Ming \eal \cite{Ming-2018} for a Johnson-Cook flow law or Liang \eal \cite{Liang-2022} for an Arrhenius type flow law.
The implementation of this VUMAT subroutine also requires the computation of the derivatives of the flow stress $\sigma^y$ with respect to the three input variables $\varepsilon^p$, $\mdot\varepsilon$ and $T$ which can quickly become relatively complex as the complexity of the flow law increases.
The choice of this flow law is therefore doubly guided by the nature of the behavior of the material on the one hand, but especially by the list of flow laws implemented natively in the FE codes.
At this time, there is not yet a flow law generic enough to cover a wide range of material behavior.

Depending on the nature of the study, the choice of this flow law is generally guided mainly by the list of flow laws available in the Finite Element code used, and very often this choice is made at the expense of the quality of the model.
For example, Zhou \eal \cite{Zhou-2019}, proposed the identification of the flow law of a GCr15 alloy for a continuous casting bloom with heavy reduction application.
In their study, they performed compression tests on GCr15 cylinders in a temperature range of $750~\celsius$ to $1300~\celsius$, strain rates of $0.001~\ps$ to $0.1~\ps$ and strains of $0$ to $0.7$.
The results of these compression tests, plotted in Figure \ref{fig:Zhou-OriginalData}, show a decrease in flow stress $\sigma^y$ with respect to temperature $T$ and a growth of $\sigma^y$ with respect to strain rate $\mdot\varepsilon$, as in most metallic materials.
\begin{figure}[!ht]
\centering
\includegraphics[width=0.9\columnwidth]{Figures/Zhou-OriginalData}
\caption{Original Data.}
\label{fig:Zhou-OriginalData}
\end{figure}
The evolution of the flow stress as a function of the plastic deformation shows the presence of a dynamic recrystallization phenomenon (DRX) within the material.
This phenomenon is an additional non-linearity of this type of material compared to other materials, mainly due to high temperatures and low strain rates, which should be taken into account when describing the material behavior.
As stated in their publication, depending on the flow model used: Johnson-Cook, Modified Zerilli-Armstrong, Arrhenius or new Modified Johnson-Cook, the fidelity of taking into account the real behavior varies widely with the complexity of the flow model which includes between $5$ parameters for the Johnson-Cook model, and up to $16$ parameters for the Arrhenius model.
Thus, and for the input data used, the two most common models: Johnson-Cook and Zerilli-Armstrong do not correctly describe the material behavior.
Only the modified Johnson-Cook and Arrhenius models are able to describe the behavior correctly.
Unfortunately, and this is not part of their study, if these last two models are satisfactory from a theoretical point of view, from a practical point of view for the user of a FE code such as Abaqus, it will be necessary to carry out a numerical implementation in a FORTRAN VUMAT subroutine of the modified Johnson-Cook flow law or the Arrhenius law as carried out by Liang \eal \cite{Liang-2022} in order to be able to use these laws for numerical simulation.
This requires a certain expertise in the development and implementation of flow laws, which is not available to all users of the Abaqus FE code.

From this observation, and from the necessity to select a flow law for a given type of material, then to identify the parameters of this flow law according to experimental tests, and finally to implement this flow law in the form of a user subroutine in FORTRAN in the Abaqus FE code, we have recently proposed in Pantalé \eal \cite{Pantale-2021} an alternative approach based on the ability of Artificial Neural Networks (ANNs) to behave as universal approximators as reported by Hornik \eal \cite{Hornik-1989}.
The underlying idea is to implement a flow law described by an ANN in the form of a FORTRAN subroutine in the Abaqus FE code.
This neural network was previously trained from the data extracted from mechanical tests of the material and can directly define the value of the flow stress $\sigma^y$ as a function of the plastic strain $\varepsilon^p$, the strain rate $\mdot\varepsilon$ and the temperature $T$.
After this learning phase based on the use of the Python library Tensorflow \cite{Abadi-2016}, the weights and biases of the neural network are transcoded into a subroutine in FORTRAN which is compiled and linked with the libraries of the Abaqus FE code in order to include the behavior of the material by allowing the computation of the flow stress $\sigma^y$ as a function of $\varepsilon^p$, $\mdot\varepsilon$ and $T$, and of its three derivatives $\partial\sigma^y/\partial\varepsilon^p$, $\partial\sigma^y/\partial\mdot\varepsilon$ and $\partial\sigma^y/\partial T$.

The structure of this paper is as follows.
Section \ref{sec:Training} deals with the presentation of a Neural Network based flow law and its training from the data reported in figure \ref{fig:Zhou-OriginalData}.
The comparison of several neural network architectures regarding accuracy and implementation complexity will be presented.
In Section \ref{sec:Use}, we will present the conversion of this neural network into a FORTRAN or subroutine for the Abaqus FE code.
Validation is based on real test cases of the use of an ANN flow law in numerical simulations and the comparison between the implementations made on Abaqus Explicit.
Finally, a conclusion and perspective section will be presented at the end.

@@

In this section, we briefly recall, as an introduction, some basic principles of Artificial Neural Networks that are relevant to this work.
The global architecture chosen to model the behavior of a material is based on a multi-layer feed-forward ANN which as proposed by Hornik \eal \cite{Hornik-1989} can be used as a universal approximator.
The architecture retained for this study concerns a neural network with two hidden layers containing a variable number of neurons on these two layers, $3$ inputs corresponding to the plastic strain $\varepsilon^p$, the strain rate $\mdot\varepsilon$ and the temperature $T$ respectively and a single output concerning the flow stress of the material $\sigma^y$.
Figure \ref{fig:ANN-scheme} shows a graphical representation of the global architecture of this neural network.
\begin{figure}[!ht]
\centering
\includegraphics[width=0.7\columnwidth]{Figures/ANN-scheme-2HL}
\caption{Global structure of the ANN flow law with two hidden layers, 3 inputs and one output.}
\label{fig:ANN-scheme}
\end{figure}
The choice of the number of neurons in the two hidden layers is free, but must be reasonable.
Indeed, the more neurons the network contains, the more it will be able to reproduce faithfully the training data, but the less it will be able to generalize to new data (the classical problem of overlearning of neural networks).
Moreover, the more neurons it contains, the more complex its mathematical structure will be, and the more computation time it will require within the routine included in the FE code.
It is therefore necessary to respect a balance between the capacity of the network to minimize errors during the learning phase, its complexity and the speed of execution once it is transcribed into the FE code.

@@

According to Figure \ref{fig:ANN-scheme}, the neural network has $3$ inputs (referred as the input vector $\overrightarrow{x}$ of the ANN) corresponding to the plastic strain $\varepsilon^p$, the strain rate $\mdot\varepsilon$ and the temperature $T$ respectively.
These inputs are first normalized within the range $[0,1]$ in order to avoid a ill-conditioning of the system as presented by many other authors in the literature \cite{Lin-2008, Lu-2011} since these three variables represent different physical data with a very different amplitudes ($0.7$ for the plastic strain, $100$ for the strain rate and $550~\celsius$ for the temperature in the case of the data reported in Figure \ref{fig:Zhou-OriginalData}).
Therefore, the three components of the input vector $\overrightarrow{x}$ are obtained from the plastic strain $\varepsilon^p$, the strain rate $\mdot\varepsilon$ and the temperature $T$ using the following expressions:
\begin{equation}
\overrightarrow{x}=
\begin{cases}
x_1 = \frac{\varepsilon^p - [\varepsilon^p]_{min}}{[\varepsilon^p]_{max} - [\varepsilon^p]_{min}}\\
x_2 = \frac{\ln(\mdot\varepsilon/\mdot\varepsilon_0)-[\ln(\mdot\varepsilon/\mdot\varepsilon_0)]_{min}}{[\ln(\mdot\varepsilon/\mdot\varepsilon_0)]_{max}-[\ln(\mdot\varepsilon/\mdot\varepsilon_0)]_{min}}\\
x_3 = \frac{T-[T]_{min}}{[T]_{max}-[T]_{min}}
\end{cases},
\label{eq:CR1}
\end{equation}
where $[~]_{min}$ and $[~]_{max}$  are the boundaries of the range of the corresponding field.
Concerning the strain rate $\mdot\varepsilon$, and taking into account that its amplitude in a real case can reach $10^5~\ps$, as proposed in Pantalé \cite{Pantale-2021}, we chose at first to substitute $\ln(\mdot\varepsilon/\mdot\varepsilon_0)$, with $\mdot\varepsilon_0$ equal to the lowest strain rate test, for the value of $\mdot\varepsilon$.
After normalization, these three input variables are introduced into the neural network and are propagated within it by the feed-forward propagation mechanism.

Conforming to the structure of the ANN reported in Figure \ref{fig:ANN-scheme} any hidden layer $k$, containing $n$ neurons, takes a weighted sum of the outputs $\overrightarrow{\hat{y}}$ of the immediately previous layer $(k-1)$, containing $m$ neurons, given by the following equation:
\begin{equation}
y_i\lay{k} = \sum_{j=1}^m w_{ij}\lay{k} \hat{y}_j^{(k-1)}+ b_i\lay{k},\label{eq:ANN1}
\end{equation}
where $y_i\lay{k}$ is the entry of the $i^{th}$ neuron of layer $k$, $\hat{y}_j\lay{k-1}$ is the output of the $j^{th}$ neuron of layer $(k-1)$, $w_{ij}\lay{k}$ is the associated weight parameter between the $i^{th}$ neuron of layer $k$ and the $j^{th}$ neuron of layer $(k-1)$ and $b_i\lay{k}$ is the associated bias of the $i^{th}$ neuron of layer $k$.
Those weights $w_{ij}$ and bias $b_i$, for each layer, are the training parameters of the ANN that we have to adjust during the training process.
For the proposed model, we have selected the Sigmoid activation function, so that, each neuron in the hidden layer $k$ provides an output value ${\hat{y}}$ from the input value $y$ of the same neuron defined by Equation (\ref{eq:ANN1}) according to the following equation:
\begin{equation}
\hat{y}=\frac{1}{1 + \e{-y}}\label{eq:ANN2}
\end{equation}

>>>>

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.
\usepackage{gensymb}
\usepackage{amsmath}
\usepackage{accents}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{subcaption}
\DeclareRobustCommand{\w}{\mbox{\large\ensuremath{\mathsf{w}}}}
\DeclareRobustCommand{\A}{\mbox{\large\ensuremath{\mathsf{A}}}}
\DeclareRobustCommand{\B}{\mbox{\large\ensuremath{\mathsf{B}}}}
\DeclareRobustCommand{\C}{\mbox{\large\ensuremath{\mathsf{C}}}}
\DeclareRobustCommand{\Dev}{\mbox{\Large\ensuremath{\mathsf{s}}}}
\DeclareRobustCommand{\Sig}{\mbox{\Large\ensuremath{\sigma}}}
\DeclareRobustCommand{\dotp}{\boldsymbol{\cdot}}
\DeclareRobustCommand{\ccirc}{\kern0.5ex\vcenter{\hbox{$\scriptstyle\circ$}}\kern0.5ex}
\DeclareRobustCommand{\e}[1]{{\rm e}^{#1}}
\DeclareRobustCommand{\lay}[1]{^{(#1)}}
\DeclareRobustCommand{\mdot}[1]{\accentset{\mbox{\bfseries .}}{#1}}
\DeclareRobustCommand{\ie}{\emph{i.e.}\@\xspace}
\DeclareRobustCommand{\eal}{et \emph{al.}\@\xspace}
\DeclareRobustCommand{\eg}{e.g.,\@\xspace}
\DeclareRobustCommand{\MSE}{\text{E}_\text{MS}}
\DeclareRobustCommand{\RMSE}{\text{E}_\text{RMS}}
\DeclareRobustCommand{\MARE}{\text{E}_\text{MAR}}
\DeclareRobustCommand{\R}{\text{R}}
\DeclareRobustCommand{\ps}{\text{s}^{-1}}
\DeclareRobustCommand{\mr}[2]{\multirow{#1}{*}{#2}}
\DeclareRobustCommand{\MPa}{\text{MPa}}

@@

Numerical methods for simulating the behavior of structures subjected to high thermomechanical loads, as in the case of high-temperature forming of metallic materials, are generally based on the use of commercial Finite Element (FE) codes such as Abaqus or laboratory codes such as DynELA \cite{Pantale-2004}.
These FE codes are based on two types of equations: conservation equations and constitutive equations.
If the first equations are generally well established on the basis of physics and mechanics, it is not the same for the second type of equations: the constitutive equations.
Thus, in a general way, the conservation equations concern the fundamental principles of physics such as the mass conservation law, the momentum law (fundamental equation) and the energy law (declined in the form of the first and second principles of thermodynamics).
By themselves, these laws are not sufficient to describe the behavior of a material or a structure subjected to thermomechanical solicitations, because the nature of the material's behavior, generally translated by means of the behavior laws, is not included in the system previously proposed.
Therefore, for each type of material, it is necessary to define behavior laws whose formulation is based on observation, in order to describe the behavior of this material under external forces.
The quality and the accuracy of the results of any numerical simulation depend on the choice of these behavior laws and on the ability of the user to identify the coefficients of these behavior laws for a given material by performing experiments under conditions close to those encountered during the real stress of the structure in service that one wishes to design.
Depending on the nature of the solicitations, these tests are based on quasi-static or dynamic tensile or compression tests, tests on thermomechanical simulators such as Gleeble \cite{Lin-2009} or impact tests using gas launchers or Hopkinson bars \cite{Kolsky-1949}.

In the context of thermomechanical simulation of forming processes, these behavior laws define the dependence of the flow stress of the material $\sigma^y$ as a function of the three input variables which are the plastic strain $\varepsilon^p$, the strain rate $\mdot\varepsilon$ and the temperature $T$ of the material, so that the general form of the flow law can be written with the following expression:
\begin{equation}
\sigma^y = f\left(\varepsilon^p,\mdot\varepsilon,T\right)\label{eq:GenFlow}
\end{equation}

These laws, due to the nature of materials and the phenomena involved (work hardening, movement of dislocations, structural hardening, phase transformations, ...) are highly non-linear and their validity is restricted to a certain range of strains $\varepsilon$, strain rates $\mdot\varepsilon$ and temperatures $T$.
From the observations made, we can define two main classes of behavior laws: the flow laws based on physics and the empirical flow laws.
From the mechanics of continuous media and experimental tests and depending on the nature of the materials used, a number of flow models have been developed in the past, including: the Johnson-Cook flow law \cite{Johnson-1983}, the Zerilli Armstrong flow law \cite{Zerilli-1987} and their respective derived forms \cite{Rule-1998, Lin-2011, Li-2013, Zhang-2015, Zhou-2020}, the Hansel-Spittle \cite{Hensel-1978} or the Arrhenius \cite{Jonas-1969} flow laws, to name only a few of the most widely used in the context of the metal forming processes at high temperature.
As an example, and because it is widely used in numerical simulation of metal forming processes, the equation that describes the Johnson-Cook flow law \cite{Johnson-1983} is given as follows:
\begin{equation}
\sigma^y=\left(A+B\varepsilon^{p^{n}}\right) \left[1+C\ln\left(\frac{\mdot\varepsilon}{\mdot\varepsilon_0}\right)\right] \left[1-\left(\frac{T-T_0}{T_m-T_0}\right)^{m}\right],\label{eq:Johnson-Cook}
\end{equation}
where $A$ is the initial elastic limit of the material, $B$ is the strain hardening coefficient, $n$ is the strain hardening exponent, $C$ and $m$ are the material constants that describe the strain rate hardening coefficient and the thermal softening coefficient, respectively.

Once the choice has been made concerning the type of flow law to be used for a given material, it is then necessary, from a set of experimental tests carried out in the laboratory under conditions close to those of the structure in service, to identify the parameters of these flow laws by machine learning methods based on approaches of minimization of the calculated experiment.
Therefore, the use of the Johnson-Cook flow law defined by Equation (\ref{eq:Johnson-Cook}) require the identification of $5$ material parameters.

The main problem that the researchers are confronted with after the phase of realization of the experimental tests concerns the choice of the flow law to use according to the observations made on these test results.
This choice of flow law is also restricted by the nature of the FE code used and the availability of such flow laws.
Thus, a user of the Abaqus FE code, will turn more particularly to a Johnson-Cook \cite{Johnson-1983} flow law insofar as it is natively implemented in this software.
The choice of another form of flow law, Zerilli-Armstrong, or Arrhenius for example, obliges the user to program himself the computation of the flow stress $\sigma^y$ of the material by means of a VUMAT subroutine in FORTRAN 77 as proposed by Gao \eal \cite{Gao-2007-FRT}, Ming \eal \cite{Ming-2018} for a Johnson-Cook flow law or Liang \eal \cite{Liang-2022} for an Arrhenius type flow law with the following expression:
\begin{equation}
\sigma^y = \frac{1}{\alpha(\varepsilon)} \ln\left\{\left(\frac{Z(\varepsilon)}{A(\varepsilon)}\right)^{1/n(\varepsilon)} + \sqrt{1 + \left(\frac{Z(\varepsilon)}{A(\varepsilon)}\right)^{2/n(\varepsilon)}}\right\}
\label{eq:ArDef}
\end{equation}
with:
\begin{equation}
Z(\varepsilon) = \mdot\varepsilon \exp{\left(\frac{Q(\varepsilon)}{RT}\right)}, \label{eq:ArZ}
\end{equation}
where $Z$ is the Zenner-Hollomon parameter \cite{Zener-1944}, $Q(\varepsilon)$ is the apparent activation energy ($\text{J~mol}^{-1}$), $R$ is the universal gas constant ($8.314~\text{J~mol}^{-1} \text{K}^{-1}$).
$Q(\varepsilon)$, $A(\varepsilon)$, $\alpha(\varepsilon)$ and $n(\varepsilon)$ are expressed as a function of the strain $\varepsilon$ by means of polynomial functions of degree $m$ (varying from $1$ to $9$), which leads to the identification of up-to $36$ material parameters.

The implementation of the flow law as a VUMAT FORTRAN subroutine requires the computation of the derivatives  $\partial\sigma^y/\partial\varepsilon^p$, $\partial\sigma^y/\partial\mdot\varepsilon$ and $\partial\sigma^y/\partial T$ of the flow stress $\sigma^y$ which can quickly become relatively complex as the complexity of the flow law increases, \ie the relative complexity of the Arrhenius flow law defined by Equations (\ref{eq:ArDef}) and (\ref{eq:ArZ}), with regard to the relative simplicity of the Johnson-Cook defined by Equation (\ref{eq:Johnson-Cook}), one can refer to the work proposed by Liang \eal \cite{Liang-2022} for details concerning this implementation using the Safe version of the Newton-Raphson method proposed by Ming \eal \cite{Ming-2018}.
The choice of the flow law to use for a given problem, is therefore doubly guided by the nature of the behavior of the material on the one hand, but a more important aspect is the list of flow laws implemented natively in the FE code we plan to use for the numerical simulation.
At this time, there is not yet a flow law generic enough to cover a wide range of material behavior that is simple to implement and use.

As we have seen in the previous paragraph, the choice of the flow law to use is generally guided mainly by the list of flow laws available in the Finite Element code used, and very often this choice is made at the expense of the quality of the model.
For example, Zhou \eal \cite{Zhou-2020}, proposed the identification of the flow law of a GCr15 alloy for a continuous casting bloom with heavy reduction application as introduced by Ji \eal \cite{Ji-2018} who have performed compression tests on this material.
In their study, Ji \eal \cite{Ji-2018} performed compression tests on GCr15 cylinders in a temperature range of $750~\celsius$ to $1300~\celsius$ in $50~\celsius$ steps, strain rates of $0.001~\ps$,  $0.01~\ps$ and $0.1~\ps$ and strains up-to $0.7$.
The results of these compression tests, plotted in Figure \ref{fig:Zhou-OriginalData}, show a decrease in flow stress $\sigma^y$ with respect to an increase of the temperature $T$ and a increase of $\sigma^y$ with respect to an increase of the strain rate $\mdot\varepsilon$, as in most metallic materials.
\begin{figure}[!ht]
\centering
\includegraphics[width=0.9\columnwidth]{Figures/Zhou-OriginalData}
\caption{Original Data.}
\label{fig:Zhou-OriginalData}
\end{figure}
The evolution of the flow stress as a function of the plastic deformation shows the presence of a dynamic recrystallization (DRX) phenomenon within the material.
This phenomenon is an additional non-linearity of this type of material compared to other materials, mainly due to high temperatures and low strain rates, which should be taken into account when describing the material behavior.
As stated in the publication of Zhou \eal \cite{Zhou-2020}, depending on the flow model used: Johnson-Cook, Modified Zerilli-Armstrong, Arrhenius or new Modified Johnson-Cook, the fidelity of taking into account the real behavior varies widely with the complexity of the flow model which includes between $5$ parameters for the Johnson-Cook model, and $16$ parameters for the Arrhenius model.
Thus, and for the input data provides by Ji \eal \cite{Ji-2018}, the two most common models: Johnson-Cook and Zerilli-Armstrong do not correctly describe the material behavior.
Only the modified Johnson-Cook and Arrhenius models are able to describe correctly the behavior of the material during the compression process.
Unfortunately, and this is not part of their study, if these last two models are satisfactory from a theoretical point of view, from a practical point of view for the user of a FE code such as Abaqus, it will be necessary to carry out a numerical implementation in a FORTRAN 77 VUMAT subroutine of the modified Johnson-Cook flow law or the Arrhenius law as carried out by few authors \cite{Liang-2022, Gao-2007-FRT, Ming-2018} in order to be able to use these laws for numerical simulation.
This requires a certain expertise in the development and implementation of flow laws, which is not available to all users of the Abaqus FE code.

From this observation, and from the necessity to select a flow law for a given type of material, then to identify the parameters of this flow law according to experimental tests, and finally to implement this flow law in the form of a user subroutine in FORTRAN in the Abaqus FE code, we have recently proposed in Pantalé \eal \cite{Pantale-2021} an alternative approach based on the ability of Artificial Neural Networks (ANNs) to behave as universal approximators as reported by Minsky \eal \cite{Minsky-1969} and Hornik \eal \cite{Hornik-1989}.
In fact, Artificial Neural Networks can solve problems that are difficult to conceptualize using traditional computational methods.
Unlike a classical approach based on a regression method, an ANN does not need to know the mathematical form of the model it seeks to reproduce, hence we do not need anymore to postulate the mathematical form of the constitutive equation to use it in a FE simulation using this kind of approach.

The implementation of ANNs for plasticity in thermomechanics has been studied and a review of the literature can be found for example in Gorji \eal \cite{Gorji-2020} concerning the use of recurrent neural networks, in Jamli \eal \cite{Jamli-2019-SNN} concerning their application in finite element analysis of metal forming processes, or in Jiao \eal \cite{Jiao-2020} concerning the applicability to meta-materials and their characterization.
A distinction must be made between ANN-based flow models (the focus of this study) and ANN-based constitutive models.
Both approaches have been studied by many researchers during the last thirty years.
Ghaboussi \cite{Ghaboussi-1991} proposed an ANN-based constitutive model for concrete under monotonic biaxial loading and cyclic uniaxial loading.
They extended their work by introducing adaptive and autoprogressive networks in \cite{Ghaboussi-1998, Ghaboussi-1998-NNA} where the architecture of the network evolves during the learning phase to better learn the complex stress-strain behavior of the materials using a global load-deflection response where the evaluation of the flow stress of the material computed by the ANN is combined with a Radial-Return algorithm.
Lin \eal \cite{Lin-2008} proposed an ANN to predict the flow stress of 42CrMo4 steel in hot compression tests on a Gleeble thermomechanical device and showed a very good correlation between the experimental results and the model predictions.
Ashtiani \eal \cite{Ashtiani-2016} compared the predictive capabilities of an ANN versus an analytical model for Johnson-Cook, Arrhenius, and strain-compensated Arrhenius laws and concluded that the neural network had better efficiency and accuracy in predicting the hot behavior of Al-Cu-Mg-Pb alloy.

The underlying idea proposed in or approach, is to implement a flow law described by a trained ANN in the form of a FORTRAN 77 subroutine in the Abaqus FE code.
This ANN was previously trained from the data extracted from mechanical tests of the material and can directly define the value of the flow stress $\sigma^y$ as a function of the plastic strain $\varepsilon^p$, the strain rate $\mdot\varepsilon$ and the temperature $T$.
After a training phase based on the use of the Python library Tensorflow \cite{Abadi-2016, Mattmann-2020}, the weights and biases of the trained neural network are transcoded into a subroutine in FORTRAN 77 which is compiled and linked with the libraries of the Abaqus FE code in order to include the behavior of the material by allowing the computation of the flow stress $\sigma^y$ as a function of $\varepsilon^p$, $\mdot\varepsilon$ and $T$, and of its three derivatives $\partial\sigma^y/\partial\varepsilon^p$, $\partial\sigma^y/\partial\mdot\varepsilon$ and $\partial\sigma^y/\partial T$.

The structure of this paper is as follows.
Section \ref{sec:Training} deals with the presentation of a Neural Network based flow law and its training from the data proposed by Ji \eal \cite{Ji-2018} and reported in Figure \ref{fig:Zhou-OriginalData}.
The comparison of several neural network architectures regarding accuracy and implementation complexity will be presented and compared.
In Section \ref{sec:Use}, we will present the transposition of this neural network into a FORTRAN 77 subroutine for the Abaqus FE code.
Validation is based on the numerical simulation Abaqus Explicit FE code of a compression test in the same configuration as the one proposed by Ji \eal \cite{Ji-2018} using four different ANN flow laws.
Finally, a conclusion and perspective Section will conclude this paper.

@@

In this section, we briefly recall, as an introduction, some basic principles of Artificial Neural Networks that are relevant to this work.
The global architecture chosen to model the behavior of a material is based on a multi-layer feed-forward ANN which as proposed by Hornik \eal \cite{Hornik-1989} can be used as a universal approximator.
The architecture retained for this study concerns a neural network with two hidden layers containing a variable number of neurons on these two layers, $3$ input nodes corresponding to the plastic strain $\varepsilon^p$, the strain rate $\mdot\varepsilon$ and the temperature $T$ respectively and a single output node for the flow stress $\sigma^y$ of the material.
Figure \ref{fig:ANN-scheme} shows a graphical representation of the global architecture of this neural network.
\begin{figure}[!ht]
\centering
\includegraphics[width=0.7\columnwidth]{Figures/ANN-scheme-2HL}
\caption{Global structure of the ANN flow law with two hidden layers, 3 input neurons ($\varepsilon^p$, $\mdot\varepsilon$, $T$) and one output neuron $\sigma^y$.}
\label{fig:ANN-scheme}
\end{figure}
The choice of the number of neurons in the two hidden layers is free, but must be reasonable.
Indeed, the more neurons the network contains, the more it will be able to reproduce faithfully the training data, but the less it will be able to generalize to new data (the classical problem of overlearning of neural networks).
Moreover, the more neurons it contains, the more complex its mathematical structure will be, and the more computation time it will require to propagate the data inside of it within the routine included in the FE code.
It is therefore necessary to respect a balance between the capacity of the network to minimize errors during the learning phase, its complexity and the computing CPU time once it is transcribed into the FE code.

@@

According to Figure \ref{fig:ANN-scheme}, the proposed neural network has $3$ inputs (referred as the input vector $\overrightarrow{x}$) corresponding to the plastic strain $\varepsilon^p$, the strain rate $\mdot\varepsilon$ and the temperature $T$ respectively.
These inputs are first normalized within the range $[0,1]$ in order to avoid a ill-conditioning of the system as presented by many other authors in the literature \cite{Lin-2008, Lu-2011} since these three variables represent different physical data with a very different amplitudes ($0.7$ for the plastic strain, $100~\ps$ for the strain rate and $550~\celsius$ for the temperature in the case of the training data reported in Figure \ref{fig:Zhou-OriginalData}).
Therefore, the three components of the input vector $\overrightarrow{x}$ are obtained from the plastic strain $\varepsilon^p$, the strain rate $\mdot\varepsilon$ and the temperature $T$ using the following expressions:
\begin{equation}
\overrightarrow{x}=
\begin{cases}
x_1 = \frac{\varepsilon^p - [\varepsilon^p]_{min}}{[\varepsilon^p]_{max} - [\varepsilon^p]_{min}}\\
x_2 = \frac{\ln(\mdot\varepsilon/\mdot\varepsilon_0)-[\ln(\mdot\varepsilon/\mdot\varepsilon_0)]_{min}}{[\ln(\mdot\varepsilon/\mdot\varepsilon_0)]_{max}-[\ln(\mdot\varepsilon/\mdot\varepsilon_0)]_{min}}\\
x_3 = \frac{T-[T]_{min}}{[T]_{max}-[T]_{min}}
\end{cases},
\label{eq:CR1}
\end{equation}
where $[~]_{min}$ and $[~]_{max}$  are the boundaries of the range of the corresponding field.
Concerning the strain rate $\mdot\varepsilon$, and taking into account that its amplitude in a real case can reach $10^5~\ps$, as proposed in Pantalé \eal \cite{Pantale-2021}, we chose at first to substitute $\ln(\mdot\varepsilon/\mdot\varepsilon_0)$, with $\mdot\varepsilon_0$ equal to the lowest strain rate test, for the value of $\mdot\varepsilon$.
After normalization, these three input variables are introduced into the neural network and are propagated within it by the feed-forward propagation mechanism.

Conforming to the structure of the ANN reported in Figure \ref{fig:ANN-scheme} any hidden layer $k$, containing $n$ neurons, takes a weighted sum of the outputs $\overrightarrow{\hat{y}}^{(k-1)}$ of the immediately previous layer $(k-1)$, containing $m$ neurons, given by the following equation:
\begin{equation}
y_i\lay{k} = \sum_{j=1}^m w_{ij}\lay{k} \hat{y}_j^{(k-1)}+ b_i\lay{k},\label{eq:ANN1}
\end{equation}
where $y_i\lay{k}$ is the entry of the $i^{th}$ neuron of layer $k$, $\hat{y}_j\lay{k-1}$ is the output of the $j^{th}$ neuron of layer $(k-1)$, $w_{ij}\lay{k}$ is the associated weight parameter between the $i^{th}$ neuron of layer $k$ and the $j^{th}$ neuron of layer $(k-1)$ and $b_i\lay{k}$ is the associated bias of the $i^{th}$ neuron of layer $k$.
Those weights $w_{ij}$ and bias $b_i$, for each layer, are the training parameters of the ANN that we have to adjust during the training process.
For the proposed model, we have selected the sigmoid activation function, so that, each neuron in the hidden layer $k$ provides an output value ${\hat{y}}$ from the input value $y$ of the same neuron defined by Equation (\ref{eq:ANN1}) according to the following equation:
\begin{equation}
\hat{y}=\frac{1}{1 + \e{-y}}\label{eq:ANN2}
\end{equation}
